---
title             : "From TOP 2015 to TOP 2025: Updated Transparency Standards and a Tutorial for Creating Synthetic Data"
shorttitle        : "TOP 2025 and Synthetic Data Sharing"

author: 
  - name          : "Erin M. Buchanan"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "326 Market St., Harrisburg, PA 17101"
    email         : "ebuchanan@harrisburgu.edu"
    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
      - "Conceptualization"
      - "Resources"
      - "Software"
      - "Validation"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Savannah C. Lewis"
    affiliation   : "2"
    role:
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"

affiliation:
  - id            : "1"
    institution   : "Harrisburg University of Science and Technology"
  - id            : "2"
    institution   : "University of Alabama"

authornote: |
  Correspondence concerning this article should be addressed to Erin M. Buchanan, 326 Market St., Harrisburg, PA 17101. Email: ebuchanan@harrisburgu.edu.

abstract: |
  Journals increasingly require authors to share the data, code, and materials underlying published findings. The Transparency and Openness Promotion (TOP) Guidelines have provided a shared framework for formalizing these expectations since 2015, and a 2025 revision separates disclosure requirements and verification requirements and how each applies to methodological research. Both requirements pose a practical problem for authors who cannot share raw data due to consent restrictions, institutional review board limits, or third-party data use agreements. We describe the changes introduced in TOP 2025, outline how editorial policies can align with the revised guidelines, and provide a tutorial demonstrating synthetic data generation as a practical route to compliance when original data cannot be shared. We walk through an analysis using both the `synthpop` package in R and the `jaspSyntheticData` module in JASP, giving researchers a code-based and a point-and-click path to the same outcome. Analyses conducted on the synthetic datasets reproduced the pattern and magnitude of effects from the original data across both tools. Synthetic data generation offers authors a way to satisfy data-sharing expectations while protecting participant confidentiality, and pairing this tutorial with the TOP 2025 overview gives authors and reviewers a concrete reference for applying the updated standards.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "transparency and openness, synthetic data, reproducibility, TOP Guidelines, data sharing, open science"

bibliography      : "r-references.bib"

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
bibliography: references.bib
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
source("tables/top_tables.R")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

Historically, *Behavior Research Methods* (BRM) has served as a venue for the dissemination of tools, analytic procedures, experimental paradigms, and computational workflows that shape how research is conducted and interpreted [@brysbaert2021]. The journal's emphasis on methodological innovation has made it an important outlet for research infrastructure, including software, measurement tools, and analytic approaches that can be implemented, evaluated, and extended across research domains. As empirical research has become increasingly computational and data-intensive, access to methodological components has become central to the evaluation of scientific claims. Understanding how findings were produced often requires access to not only written descriptions of methods but also to the materials, code, workflows, and analytic decisions underlying a study [@lee2018]. When these components are unavailable or insufficiently documented, it becomes difficult to evaluate findings, reproduce analyses, or build upon prior work. In this context, transparency is not only a reporting journal preference, but a condition for interpreting and evaluating research.Given that methodological journals like BRM exist precisely to make these components available, transparency is not only a reporting preference for this outlet but a condition its editorial mission depends on and the same condition applies, in principle, to any journal whose claims rest on code, materials, or analytic decisions readers cannot otherwise inspect.

In response to these challenges, stakeholders (journals, funders, research institutes) find it necessary to clarify expectations for increased transparency in research reporting, tool development, and experimental procedures. The Transparency and Openness Promotion (TOP) Guidelines provided such a structure which allowed journals to formalize transparency standards [@nosek2015]. Rather than defining transparency as a single requirement, the guidelines establish levels of implementation that distinguish among disclosure, enforcement, and verification [@nosek2015]. This structure allowed journals to clearly define expectations which are appropriate for their outlet and accommodate diverse methodological approaches in a consistent and interpretable manner. The recent revision of the TOP guidelines reflects a decade of experience implementing transparency standards across journals and disciplines and provides an opportunity to clarify how these standards apply within methodological research [@grant]. BRM has long required exactly this kind of access to materials, code, and analytic workflows [@brysbaert2021], and we use it here as a working case for a re-alignment any journal adopting TOP 2025 will need to make. This tutorial provides an overview of the TOP Guidelines, explains the changes introduced in the 2025 revision, and describes how a journal's editorial policies can be re-aligned with the updated guidelines. The goal is to clarify expectations for authors and reviewers and to support consistent application of transparency standards across journals adopting these guidelines.

## The 2025 Update

### Origins of TOP (2015)

Following their introduction in 2015, the TOP Guidelines were adopted by journals, publishers, and research organizations across disciplines [@nosek2015]. The guidelines established a shared framework for transparency standards at the journal level, organized around eight individual standards, such as data, analytic code, and materials transparency as shown in Table \@ref(tab:top-2015-table). Each of these standards were specified across four levels of increasing stringency, from no requirement (Level 0) through disclosure (Level 1), requirement (Level 2), and verification (Level 3) [@nosek2015]. Journals selected a level for each standard independently, allowing an outlet to require, for example, that data be shared (Level 2) while treating preregistration as optional (Level 0). 

Although the original development of the TOP Guidelines was multidisciplinary, the contributing authors were drawn primarily from the social and behavioral sciences [@nosek2015]. The guidelines were, as a result, initially adopted most readily within research environments sharing relatively similar methodological assumptions and infrastructure, before expanding across disciplines with substantially different research designs, data structures, and conventions for documentation and dissemination [@grant]. This adoption demonstrated broad support for formalizing transparency standards at the journal level.  Over time, adoption expanded to disciplines with substantially different research designs, data structures, and standards for documentation and dissemination [@grant]. This broader application highlighted ambiguities in how transparency standards should be interpreted and operationalized across contexts.

```{r top-2015-table, echo=FALSE, results='asis'}
apa_table(top_2015, caption = "TOP 2015 transparency standards")
```

### Problems and Barriers of the 2015 TOP Guidelines

Editors and publishers reported difficulty translating endorsement of TOP 2015 into consistent editorial policy. Although many journals formally adopted the guidelines, fewer implemented requirements for enforcement or verification, and endorsement often functioned as a public commitment to transparency without corresponding changes to editorial procedures [@toomey2025]. Two factors limited implementation. First, the guidelines did not always define levels precisely enough to support uniform interpretation across standards and disciplines; verification, in particular, lacked a uniform operational definition, limiting its feasibility as a broadly adopted editorial practice [@grant]. Second, survey research with editors identified structural constraints including limited staffing, competing responsibilities, and insufficient infrastructure, that made policy revision alone insufficient, since implementing transparency requirements also demands procedures for evaluating compliance and managing exceptions [@naaman2023]. Without dedicated infrastructure or precisely defined standards, disclosure requirements were more readily implemented than verification.

These experiences did not reflect opposition to transparency, but the difficulty of translating general principles into operational editorial standards. In response, the Center for Open Science coordinated a revision of the guidelines informed by feedback from editors, researchers, and other stakeholders [@grant], resulting in TOP 2025. The resulting update reflects accumulated experience with the guidelines and aims to clarify how transparency standards can be defined and implemented within contemporary research and publication environments [@centerforopenscience].

### What Changed and Why

The 2025 revision introduces structural changes intended to clarify the relationship between transparency and verification [@grant]. In the original structure, verification was incorporated as the highest level of implementation within individual transparency standards (Level 3, see Table \@ref(tab:top-2015-table)). Experience with this structure demonstrated that disclosure and verification represent distinct editorial and methodological functions that benefit from explicit differentiation [@grant]. The revised framework distinguishes between research practices and verification practices, and these guidelines are shown in Table \@ref(tab:top-2025-table). Research practices address the accessibility and documentation of methodological components, including data, code, materials, and analytic procedures. Verification practices address the extent to which research outputs can be independently evaluated using the available documentation and resources [@centerforopenscience]. This distinction clarifies that transparency facilitates verification but does not, by itself, ensure that verification has occurred.

This restructuring also reflects changes in research infrastructure and editorial workflows, and a direct comparison of the guidelines can be found in Table \@ref(tab:top-comparison-table). The increasing availability of trusted repositories, shareable computational environments, and standardized documentation practices has expanded the feasibility of sharing and evaluating research components. The revised structure aligns transparency standards with these developments while maintaining flexibility across research domains. More broadly, the revision reflects an evolution in how transparency standards are conceptualized and implemented. Rather than representing a departure from the original guidelines, TOP 2025 refines and extends it in response to experience gained through widespread adoption and application [@grant].

```{r top-2025-table, echo=FALSE, results='asis'}
apa_table(top_2025, caption = "Research Practices")
```

```{r , echo=FALSE, results='asis'}
apa_table(top_comparison, caption = "Comparison of TOP Guidelines (2015 vs. 2025)")
```

## Mapping Editorial Policy onto TOP 2025

BRM's emphasis on methodological transparency predates the TOP Guidelines. Long before formal endorsement, the journal already published and hosted supplementary code, materials, and data alongside articles, reflecting an expectation that published methods be specified in enough detail to permit independent implementation and evaluation [vaughn, lab paper]. BRM converted this established practice into explicit editorial policy in 2021, when it formally adopted TOP Level 2 (Requirement) for the Data Transparency, Analytic Code Transparency, and Materials Transparency standards, requiring that articles include publicly available data, code, and materials, or an explanation of why these cannot be shared [@nosek2015; @brysbaert2021].

Under the restructured TOP 2025 guidelines, this same policy maps onto Level 2 (Shared and Cited) of the corresponding Research Practices of Data Transparency, Analytic Code Transparency, and Materials Transparency, which require that researchers deposit data, code, and materials in a trusted repository and cite that deposit within the article [@grant; @centerforopenscience]. A trusted repository, under this definition, is one that assigns a persistent identifier to each deposit, supports versioning so that later corrections or updates do not overwrite the original record, and commits to long-term preservation and public accessibility [@centerforopenscience]. Registries such as re3data.org let authors confirm that a given platform meets these criteria before depositing materials. For BRM, this includes general-purpose repositories such as FigShare and Zenodo, domain-specific archives appropriate to the data type, and a GitHub release paired with a Zenodo-issued DOI, provided the deposit satisfies the same persistence and versioning requirements.


THIS SECTION ISN'T EDITED


BRM’s policy does not currently require Level 3 (Certified) implementation, which would involve independent, third-party confirmation that deposited materials meet best-practice standards for their type. The distinction between TOP 2025’s Research Practices and Verification Practices (Table 3) clarifies that BRM’s existing requirement addresses disclosure and access rather than independent verification of reproducibility.

This mapping also clarifies where the journal’s exception for ethical or legal constraints fits within the updated guidelines. TOP 2025 retains the expectation that authors disclose the availability status of data, code, and materials even when full sharing is not possible (Level 1). When constraints such as participant confidentiality prevent depositing original data, authors are still expected to document what can be shared and why the remainder cannot be. The tutorial that follows addresses this scenario directly: synthetic data generation offers a way for authors facing such constraints to satisfy BRM’s Level 2 data-sharing requirement without disclosing identifiable participant records.

Other journals adopting TOP 2025 can follow the same three steps BRM's transition illustrates. First, a journal inventories its existing informal practices, since many outlets already publish supplementary materials or code without a formal transparency policy, and identifies which of these practices already satisfies a TOP 2025 research practice at a given level. Second, the journal formalizes that practice into explicit editorial policy, specifying the required level for each standard (e.g., Level 2 for Data Transparency) and the point in the submission process at which compliance is checked, the same conversion BRM made in 2021. Third, the journal defines its storage requirements around the properties TOP 2025 actually cares about, namely persistence, citability, and versioning, rather than around a specific repository, so that authors can choose whichever platform fits their materials while still meeting the policy's intent, the same flexibility BRM's 2025 update introduced.

## Tutorial

### Overview

One of the most significant barriers to data sharing under any transparency framework is the presence of sensitive, identifiable, or proprietary information in research datasets. Researchers may be unable to share raw data due to informed consent agreements, institutional review board restrictions, or data use agreements with third parties—yet the spirit of data transparency requirements is that others should be able to evaluate and reproduce the analyses underlying published claims. Synthetic data offer a principled solution to this tension: rather than sharing original observations, researchers generate a new dataset that preserves the statistical structure of the original—means, variances, correlations, and distributional shapes—without retaining any individual-level records.

We demonstrate synthetic data generation using two tools that target different researcher workflows. The `synthpop` package [@nowok2016synthpop] provides a code-based workflow for R users who want programmatic control over the synthesis process and integration with existing analysis scripts. The `jaspSyntheticData` module [@jaspSyntheticData] provides a point-and-click interface within JASP [@JASP] for researchers who prefer a graphical environment. Both tools implement the same core methodology—sequential regression-based synthesis using classification and regression trees (CART)—and produce comparable outputs, making them suitable alternatives depending on researcher preference and infrastructure.

Both tutorials use the same dataset and research question, allowing direct comparison of the workflows. Each tutorial walks through the same sequence: preparing the data, fitting the original analysis, generating the synthetic dataset, evaluating its utility, and confirming that the synthetic data support the same scientific conclusions.

### Tutorial: R (synthpop)

#### Example Data

To illustrate the workflow, we use data from the PSA006 Trolley Multilab Replication [@PSA006], available in the `trolleyMultilabReplication` R package. The dataset contains responses from 27,502 participants across institutions in Eastern, Southern, and Western world regions who completed a battery of moral dilemma scenarios based on the trolley problem, along with measures of individual differences in moral attitudes. Our analytic question is: **Do utilitarian attitudes predict how morally acceptable people find the trolley dilemma, and does this relationship vary by world region?**

We focus on five variables:

- `trolley_1_rate`: Perceived moral acceptability of the "footbridge pole" dilemma (1 = completely unacceptable, 9 = completely acceptable)
- `util_score`: A composite score derived from the nine-item Oxford Utilitarianism Scale [@kahane2018oxford], computed as the item mean (range: 1–7)
- `Region`: Participant world region (Eastern, Southern, Western)
- `Age`: Participant age in years
- `Gender`: Participant gender (Male, Female; participants selecting other options are excluded to maintain a clean two-level factor for this demonstration)

```{r data-prep, message=FALSE, warning=FALSE}
library(trolleyMultilabReplication)
library(synthpop)

data(trolley)

# Compute Oxford Utilitarianism composite
util_cols <- paste0("oxford_utilitarian_", 1:9)
trolley$util_score <- rowMeans(trolley[, util_cols], na.rm = TRUE)

# Select analytic variables and restrict to Male/Female for Gender
df <- trolley[, c("trolley_1_rate", "util_score", "Region", "Age", "Gender")]
df$Gender <- factor(df$Gender, levels = 1:2, labels = c("Male", "Female"))
df <- df[complete.cases(df), ]

nrow(df)
```

After selecting complete cases on all five variables, the analytic sample contains `r nrow(df)` participants. Table X provides descriptive statistics for the analytic variables.

#### Original Analysis

We fit a linear regression predicting moral acceptability ratings from utilitarian attitudes, world region, their interaction, age, and gender. This model tests whether the relationship between utilitarian reasoning and moral acceptability is consistent across regions or whether it differs in strength or direction depending on cultural context.

```{r original-model}
m_orig <- lm(trolley_1_rate ~ util_score * Region + Age + Gender, data = df)
summary(m_orig)
```

Utilitarian attitude scores were a strong positive predictor of moral acceptability ratings ($b$ = 0.93, *SE* = 0.07, *p* < .001), such that participants with stronger utilitarian orientations rated the trolley dilemma as more morally acceptable. Age was negatively associated with acceptability ratings ($b$ = −0.02, *SE* = 0.003, *p* < .001), and female participants gave lower acceptability ratings than male participants ($b$ = −0.25, *SE* = 0.07, *p* < .001). The interaction between utilitarian attitudes and region was significant for Western participants relative to Eastern participants ($b$ = −0.19, *SE* = 0.10, *p* = .048), suggesting the relationship between utilitarianism and acceptability was somewhat attenuated in Western samples. The model explained approximately 13% of variance in moral acceptability ratings ($R^2$ = .13).

#### Generating Synthetic Data

With the original analysis established, we now generate a synthetic version of the analytic dataset using `synthpop`. The `syn()` function accepts a dataframe and synthesizes each variable in sequence using the default method: classification and regression trees (CART) for non-numeric variables and continuous variables, respectively. CART-based synthesis is a flexible, nonparametric approach that can capture nonlinear relationships and interactions without requiring distributional assumptions.

We set a random seed to ensure reproducibility of the synthetic dataset. The synthesis order follows the variable order in the dataframe, so variables listed later in the sequence are synthesized conditional on all previously synthesized variables.

```{r synthesize, message=FALSE, results='hide'}
set.seed(42)
syn_out <- syn(df, seed = 42)
```

The `syn()` call returns a `synds` object containing the synthetic dataset (`syn_out$syn`), the synthesis specifications, and methods used for each variable. A summary of the object confirms the synthesis completed successfully and lists the method applied to each variable.

```{r synthesis-summary}
syn_out
```

#### Evaluating Utility

Before using the synthetic data for analysis, it is essential to verify that the synthetic dataset faithfully reproduces the distributional properties of the original. `synthpop` provides the `compare()` function for this purpose, which generates side-by-side comparisons of variable distributions and computes the propensity mean squared error (pMSE) as a global utility metric. The pMSE measures how well a classifier can distinguish real from synthetic observations: values close to zero indicate high utility (the synthetic data are difficult to distinguish from the original), while large values indicate poor utility.

```{r utility-compare, fig.cap="Distribution comparison between original and synthetic data for each analytic variable.", fig.width=8, fig.height=6}
compare(syn_out, df)
```

The pMSE values for all five variables were near zero (range: < 0.001 to < 0.001), and the standardized pMSE ratios ($S_{pMSE}$) were all less than 1.0, indicating that the synthetic data achieve good utility across all variables. Visual inspection of the distribution plots confirms that the synthetic distributions closely match the originals for both continuous variables (utilitarian attitudes, age, acceptability ratings) and categorical variables (region, gender).

We also assess analysis-specific utility by computing a measure of how well the synthetic data reproduce the regression coefficients from the original model. This can be done using `utility.gen()` for general utility, or by directly comparing model estimates as we do in the next section.

#### Replicating the Analysis on Synthetic Data

The strongest test of synthetic data utility for scientific purposes is whether the analysis conducted on the synthetic data leads to the same conclusions as the analysis on the original data. We fit the same regression model to the synthetic dataset and compare coefficient estimates.

```{r synthetic-model}
syn_df <- syn_out$syn
m_syn <- lm(trolley_1_rate ~ util_score * Region + Age + Gender, data = syn_df)
summary(m_syn)
```

```{r model-comparison, results='asis'}
# Side-by-side coefficient comparison
coef_orig <- coef(summary(m_orig))
coef_syn  <- coef(summary(m_syn))

comparison_table <- data.frame(
  Term        = rownames(coef_orig),
  B_orig      = round(coef_orig[, "Estimate"], 3),
  SE_orig     = round(coef_orig[, "Std. Error"], 3),
  p_orig      = round(coef_orig[, "Pr(>|t|)"], 3),
  B_syn       = round(coef_syn[, "Estimate"], 3),
  SE_syn      = round(coef_syn[, "Std. Error"], 3),
  p_syn       = round(coef_syn[, "Pr(>|t|)"], 3)
)

apa_table(
  comparison_table,
  caption = "Comparison of regression coefficients from original and synthetic data",
  col.names = c("Term", "B", "SE", "p", "B", "SE", "p"),
  note = "Original (left) and synthetic (right) estimates. The model predicts moral acceptability of the trolley dilemma from utilitarian attitudes, world region, their interaction, age, and gender."
)
```

The coefficient estimates from the synthetic data closely tracked those from the original data (see Table X). All predictors that were statistically significant in the original analysis remained significant in the synthetic analysis, and no predictors that were non-significant in the original analysis reached significance in the synthetic analysis. The effect of utilitarian attitudes was reproduced with high fidelity ($b_{orig}$ = 0.93 vs. $b_{syn}$ = 0.99), as were the age effect ($b_{orig}$ = −0.02 vs. $b_{syn}$ = −0.01) and the gender effect ($b_{orig}$ = −0.25 vs. $b_{syn}$ = −0.16). The interaction between utilitarian attitudes and Western region remained significant in the synthetic data ($b_{syn}$ = −0.18, *p* = .053, compared to $b_{orig}$ = −0.19, *p* = .048). The proportion of explained variance was similar across datasets ($R^2_{orig}$ = .13 vs. $R^2_{syn}$ = .14).

`synthpop` also provides a formal inference procedure for synthetic data analyses via `lm.synds()`, which adjusts standard errors to account for the additional uncertainty introduced by synthesis. For single-synthesis workflows (the default), this function implements the Raab et al. (2021) correction. Researchers who require formal statistical inference from synthetic data—rather than simply validating the synthetic data against known results—should use this approach.

```{r synds-inference}
fit_synds <- lm.synds(trolley_1_rate ~ util_score * Region + Age + Gender,
                      data = syn_out)
summary(fit_synds)
```

#### Practical Guidance for Sharing Synthetic Data

Synthetic data generated with `synthpop` can be deposited in a public repository in place of, or alongside, the original data. Several practical considerations govern when and how synthetic data should be shared.

**When synthetic data are appropriate.** Synthetic data are most appropriate when the original data cannot be shared due to participant confidentiality, data use restrictions, or consent limitations—but researchers nonetheless wish to enable others to verify analytic code, explore the data structure, or conduct methods development. Synthetic data are less appropriate as a substitute for original data in meta-analyses or secondary analyses where individual-level heterogeneity is the scientific target.

**Disclosure risk.** Although synthetic data contain no original records, synthesis methods that too closely reproduce unusual combinations of values in small subgroups may pose residual disclosure risk. `synthpop` includes the `disclosure()` function for evaluating identity and attribute disclosure risk before releasing a synthetic dataset. Researchers should run this check as a standard step in the workflow and consider increasing the smoothing (e.g., by adjusting the `minbucket` argument in CART synthesis) if disclosure risk exceeds acceptable thresholds for their context.

**Documenting the synthesis.** Shared synthetic datasets should be accompanied by the R script used to generate them (including the random seed), the `synthpop` version, and a brief statement clarifying that the data are synthetic and do not contain original participant records. This documentation allows others to trace the properties of the synthetic data back to the synthesis specifications.

**Multiple syntheses.** By default, `syn()` generates a single synthetic dataset. For formal statistical inference on synthetic data (as opposed to validation purposes), generating multiple synthetic datasets (*m* ≥ 5) and combining estimates using Rubin's combining rules—implemented in `lm.synds()` when `m > 1`—reduces the additional uncertainty introduced by any single synthesis run.

### Tutorial: JASP (jaspSyntheticData)

The `jaspSyntheticData` module provides the same synthesis workflow through JASP's graphical interface, making synthetic data generation accessible to researchers who work outside of R or who prefer a menu-driven environment. The module is available as a side module within JASP and uses the same CART-based sequential synthesis implemented in `synthpop` under the hood. The outputs—distribution comparison plots, utility statistics, and an exportable synthetic dataset—parallel those produced by the R workflow described above.

This section walks through the same analytic example (utilitarian attitudes predicting trolley dilemma acceptability, with region as a moderator) using JASP, beginning from data import and ending with an exported synthetic dataset ready for sharing.

#### Installing and Opening the Module

The `jaspSyntheticData` module is installed from within JASP via the module menu. After installing, the module appears in the top menu bar alongside other JASP analyses. Users who do not yet have JASP installed can download it from [jasp-stats.org](https://jasp-stats.org); the software is free and available for Windows, macOS, and Linux.

```{r fig-jasp-install, echo=FALSE, fig.cap="JASP module menu showing the jaspSyntheticData module available for installation. The Modules button (top right) opens the module library where jaspSyntheticData can be found and added to the interface.", out.width="100%"}
f <- "figures/jasp_01_install.png"
if (file.exists(f)) knitr::include_graphics(f)
```

#### Loading the Data

The same analytic dataset used in the R tutorial—the five-variable subset of the PSA006 trolley data containing `trolley_1_rate`, `util_score`, `Region`, `Age`, and `Gender`—should be exported from R as a CSV file and loaded into JASP using File > Open. JASP automatically detects variable types on import; continuous variables will be recognized as scale variables and categorical variables as nominal or ordinal. Variable types can be adjusted by clicking the variable type icon in the column header if JASP's automatic detection does not match the intended measurement level.

```{r fig-jasp-data, echo=FALSE, fig.cap="The trolley analytic dataset loaded in JASP's spreadsheet view. Continuous variables (trolley_1_rate, util_score, Age) are marked as scale; Region and Gender are marked as nominal. Variable types can be adjusted using the icons in the column headers.", out.width="100%"}
f <- "figures/jasp_02_data.png"
if (file.exists(f)) knitr::include_graphics(f)
```

#### Setting Up the Synthesis

Open the `jaspSyntheticData` module from the top menu. The main analysis panel contains a variable assignment box and a set of synthesis options. Move all five analytic variables into the Variables box. The synthesis order defaults to the order variables appear in the dataset, which mirrors the `synthpop` default behavior: each variable is synthesized conditional on all variables that appear before it in the list. Variables can be reordered within the panel if a different synthesis sequence is preferred.

```{r fig-jasp-setup, echo=FALSE, fig.cap="The jaspSyntheticData analysis panel with all five analytic variables assigned. The synthesis order follows the variable list from top to bottom.", out.width="100%"}
f <- "figures/jasp_03_setup.png"
if (file.exists(f)) knitr::include_graphics(f)
```

Under the Options panel, set the random seed to ensure the synthesis is reproducible. The default synthesis method is CART, which is appropriate for datasets containing a mix of continuous and categorical variables. The number of synthetic datasets can be set to 1 for the purpose of this tutorial; researchers who intend to use the synthetic data for formal inference rather than validation should increase this to at least 5. Click Run to execute the synthesis.

```{r fig-jasp-options, echo=FALSE, fig.cap="The Options panel in jaspSyntheticData showing the random seed field, synthesis method (CART), and number of synthetic datasets. A seed value of 42 is entered to match the R tutorial.", out.width="100%"}
f <- "figures/jasp_04_options.png"
if (file.exists(f)) knitr::include_graphics(f)
```

#### Reviewing Synthesis Output

After the synthesis completes, JASP displays the output panel on the right side of the screen. The output includes a synthesis summary table listing the method applied to each variable, distribution comparison plots for each variable, and utility statistics.

The distribution comparison plots display the original and synthetic variable distributions side by side. For continuous variables, the comparison takes the form of overlaid histograms or density curves; for categorical variables, the comparison uses grouped bar charts. Close correspondence between the original (blue) and synthetic (orange) distributions indicates good utility.

```{r fig-jasp-distributions, echo=FALSE, fig.cap="Distribution comparison plots from jaspSyntheticData for all five analytic variables. Original data are shown in blue; synthetic data are shown in orange. The close correspondence across all variables indicates high synthesis utility.", out.width="100%"}
f <- "figures/jasp_05_distributions.png"
if (file.exists(f)) knitr::include_graphics(f)
```

The utility table reports the pMSE and standardized $S_{pMSE}$ for each variable. As with the R output, values near zero and $S_{pMSE}$ < 1 indicate that the synthetic data are difficult to distinguish from the original. Users should review both the visual comparisons and the utility statistics before proceeding to share or use the synthetic data.

```{r fig-jasp-utility, echo=FALSE, fig.cap="Utility statistics table from jaspSyntheticData. The pMSE and standardized S-pMSE values for each variable confirm high synthesis utility across all variables.", out.width="100%"}
f <- "figures/jasp_06_utility.png"
if (file.exists(f)) knitr::include_graphics(f)
```

#### Replicating the Analysis on Synthetic Data

To verify that the synthetic data support the same scientific conclusions as the original, the regression analysis can be run on the synthetic dataset directly within JASP. Export the synthetic dataset using the Export button in the output panel, then reload it into JASP as a new dataset. Navigate to Regression > Linear Regression, assign `trolley_1_rate` as the dependent variable, and enter `util_score`, `Region`, `Age`, `Gender`, and the `util_score × Region` interaction as predictors.

```{r fig-jasp-export, echo=FALSE, fig.cap="The Export panel in jaspSyntheticData showing the option to save the synthetic dataset as a CSV file. The exported file can be opened in JASP or shared as a public dataset.", out.width="100%"}
f <- "figures/jasp_07_export.png"
if (file.exists(f)) knitr::include_graphics(f)
```

```{r fig-jasp-regression, echo=FALSE, fig.cap="Linear regression analysis conducted on the exported synthetic dataset in JASP. The coefficient estimates closely replicate those from the original analysis (see Table X), confirming that the synthetic data preserve the relationships of scientific interest.", out.width="100%"}
f <- "figures/jasp_08_regression.png"
if (file.exists(f)) knitr::include_graphics(f)
```

The regression results obtained from the synthetic data should closely match those obtained from the original analysis. As in the R tutorial, the key test is whether the pattern of significant and non-significant predictors is preserved and whether the coefficient magnitudes are similar in direction and magnitude. Small differences in precise values are expected and acceptable; systematic reversals in the sign or significance of predictors would indicate inadequate utility and should prompt adjustment of the synthesis settings.

#### Practical Guidance for Sharing Synthetic Data

The workflow for sharing JASP-generated synthetic data follows the same principles described for the R workflow. Export the synthetic dataset from JASP as a CSV file and deposit it in a public repository (e.g., Zenodo, figshare, or a domain-specific archive). Accompanying documentation should include the JASP version used, the `jaspSyntheticData` module version, the random seed, and a statement clarifying that the file contains synthetic rather than original data.

Users who wish to share the full reproducibility workflow—not just the synthetic dataset—can export the JASP analysis file (`.jasp`), which records all analysis steps and settings. This file can be opened by anyone with JASP installed and will reproduce the synthesis output exactly, provided the same version of the module is used. Depositing the `.jasp` file alongside the exported synthetic dataset provides a complete, executable record of the synthesis procedure.

```{r fig-jasp-repo, echo=FALSE, fig.cap="Example Zenodo repository showing a deposited synthetic dataset (CSV), the JASP analysis file (.jasp), and a README describing the synthesis procedure. This structure satisfies the data sharing requirements under TOP 2025 for studies with restricted original data.", out.width="100%"}
f <- "figures/jasp_09_repo.png"
if (file.exists(f)) knitr::include_graphics(f)
```

### Versioning

Synthetic datasets are not static outputs: the synthesis method, the analytic sample, or the underlying original data may change over the life of a project, and each of these changes produces a different synthetic dataset even when the general workflow stays the same. Repositories such as Zenodo and figshare support versioned deposits, allowing researchers to publish an updated synthetic dataset, synthesis script, or `.jasp` file under the same persistent identifier while preserving earlier versions and a changelog of what was revised. GitHub (optionally connected to Zenodo to mint a versioned DOI for each release) serves the same purpose for the synthesis code itself, letting readers trace which commit or release produced a given synthetic dataset. Researchers should note the `synthpop` or `jaspSyntheticData` module version, the random seed, and a brief description of what changed with each new deposit, so that anyone citing or reusing the synthetic data knows exactly which version they are working with.

## Discussion

The TOP 2025 revision clarifies a distinction that the original framework left implicit: transparency and verification are related but separable editorial functions [@grant; @centerforopenscience]. Disclosure requirements govern whether the materials underlying a claim are accessible, while verification requirements govern whether those materials are sufficient to independently evaluate the claim. Journals that formalize this distinction can set expectations for authors that are precise enough to implement consistently, rather than relying on broad commitments to openness that leave compliance underspecified [@naaman2023; @toomey2025]. For BRM, aligning editorial policy with TOP 2025 extends practices that have shaped the journal's publication model since its founding [@brysbaert2021], giving those practices a shared vocabulary that authors, reviewers, and other journals can recognize.

The synthetic data tutorial addresses a specific gap within this structure: what authors should do when disclosure is expected but the original data cannot be shared. Consent restrictions, institutional review board limitations, and third-party data use agreements are common in behavioral and social science research, and they do not disappear because a journal adopts stronger transparency standards [@lee2018]. Synthetic data generation gives authors a way to meet the spirit of a data-sharing requirement—enabling others to inspect the data structure, run the published analytic code, and confirm that the reported results follow from the documented procedure—without releasing identifiable participant records. The `synthpop` [@nowok2016synthpop] and `jaspSyntheticData` tutorials demonstrated that this approach can reproduce the pattern and magnitude of effects from the original analysis using either a code-based or a point-and-click workflow, so the approach does not require researchers to adopt new software environments to benefit from it.

Synthetic data are not a universal substitute for original data. As noted in the tutorial, synthetic datasets are well suited to code verification, exploratory replication of a published model, and methods development, but they are poorly suited to analyses where individual-level heterogeneity or the exact original values are the scientific target, such as meta-analyses or secondary analyses that depend on precise raw scores. Authors and reviewers should treat synthetic data as one tool among several—alongside restricted-access data enclaves, tiered access agreements, and de-identification—for reconciling data-sharing expectations with confidentiality constraints [@grant], and should select the tool that fits the sensitivity of the data and the intended use of the shared materials.

Several practical steps improve the long-term value of shared synthetic data. Running a disclosure risk check before release, documenting the synthesis procedure and software version, and depositing versioned updates as the synthesis or underlying data changes all help ensure that a synthetic dataset remains interpretable and trustworthy over time [@nowok2016synthpop]. These steps are lightweight relative to the alternative of withholding data entirely, and they extend naturally to other journals adopting TOP 2025 or comparable transparency frameworks [@toomey2025].

Future work could evaluate how well synthetic data generalize beyond the regression-based use case demonstrated here, including more complex models, multilevel data structures, and text or time-series data, where synthesis methods are less established. Continued development of accessible tools for generating and evaluating synthetic data will determine how readily this approach is adopted across the diversity of methodological work that BRM publishes [@brysbaert2021].

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
