---
title             : "The title"
shorttitle        : "Title"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  One or two sentences to put the results into a more **general context**.
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : "r-references.bib"

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
bibliography: references.bib
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
source("tables/top_tables.R")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

Historically, Behavior Research Methods (BRM) has served as a venue for the dissemination of tools, analytic procedures, experimental paradigms, and computational workflows that shape how research is conducted and interpreted @brysbaert2021. The journal's emphasis on methodological innovation has made it an important outlet for research infrastructure, including software, measurement tools, and analytic approaches that can be implemented, evaluated, and extended across research domains.

As empirical research has become increasingly computational and data-intensive, access to methodological components has become central to the evaluation of scientific claims. Understanding how findings were produced often requires access to not only written descriptions of methods but also to the materials, code, workflows, and analytic decisions underlying a study @lee2018. When these components are unavailable or insufficiently documented, it becomes difficult to evaluate findings, reproduce analyses, or build upon prior work. In this context, transparency is not only a reporting journal preference, but a condition for interpreting and evaluating research.

In response to these challenges, stakeholders (journals, funders, research institutes) find it necessary to clarify expectations for increased transparency in research reporting, tool development, and experimental procedures. The Transparency and Openness Promotion (TOP) Guidelines provided such a structure which allowed journals to formalize transparency standards @nosek2015. Rather than defining transparency as a single requirement, the guidelines establish levels of implementation that distinguish among disclosure, enforcement, and verification @nosek2015. This structure allowed journals to clearly define expectations which are appropriate for their outlet and accommodate diverse methodological approaches in a consistent and interpretable manner.

These developments are directly relevant to BRM’s scope and publication model. The journal publishes methodological contributions intended for implementation and reuse, transparency and accessibility are integral to the interpretability and utility of published work. The recent revision of the TOP guidelines reflects a decade of experience implementing transparency standards across journals and disciplines and provides an opportunity to clarify how these standards apply within methodological research @grant. This tutorial provides an overview of the TOP Guidelines, explains the changes introduced in the 2025 revision, and describes how BRM’s editorial policies align with the updated framework. The goal is to clarify expectations for authors and reviewers and to support consistent application of transparency standards within the journal.

## The 2025 Update

### Origins of TOP (2015)

Following their introduction in 2015, the TOP Guidelines were adopted by journals, publishers, and research organizations across disciplines @nosek2015. This adoption demonstrated broad support for formalizing transparency standards at the journal level. At the same time, implementation across diverse editorial and methodological contexts revealed areas where the original framework lacked sufficient clarity to support consistent interpretation and application @grant.

Although the original development of the TOP Guidelines was multidisciplinary, the contributing authors are primarily drawn from the social and behavioral sciences @nosek2015. As a result, the framework was initially implemented in research environments with relatively similar methodological assumptions and infrastructure. Over time, adoption expanded to disciplines with substantially different research designs, data structures, and standards for documentation and dissemination @grant. This broader application highlighted ambiguities in how transparency standards should be interpreted and operationalized across contexts.

Editors and publishers reported particular challenges in translating endorsement of the guidelines into consistent editorial policy. While the framework provided a shared vocabulary for transparency, distinctions among levels of implementation were not always sufficiently precise to support uniform interpretation. These challenges reflected the difficulty of defining general standards that could accommodate variation in research practice while remaining operationally meaningful @naaman2023.

In response to these developments, the Center for Open Science coordinated a revision of the guidelines informed by feedback from editors, researchers, and other stakeholders @grant. The resulting update reflects accumulated experience with the framework and aims to clarify how transparency standards can be defined and implemented within contemporary research and publication environments @centerforopenscience.

```{r top-2015-table, echo=FALSE, results='asis'}
apa_table(top_2015, caption = "TOP 2015 transparency standards")
```

### What Changed and Why

The 2025 revision introduces structural changes intended to clarify the relationship between transparency and verification @grant. In the original framework, verification was incorporated as the highest level of implementation within individual transparency standards. Experience with this structure demonstrated that disclosure and verification represent distinct editorial and methodological functions that benefit from explicit differentiation @grant.

The revised framework distinguishes between research practices and verification practices. Research practices address the accessibility and documentation of methodological components, including data, code, materials, and analytic procedures. Verification practices address the extent to which research outputs can be independently evaluated using the available documentation and resources @centerforopenscience. This distinction clarifies that transparency facilitates verification but does not, by itself, ensure that verification has occurred.

This restructuring also reflects changes in research infrastructure and editorial workflows. The increasing availability of trusted repositories, executable computational environments, and standardized documentation practices has expanded the feasibility of sharing and evaluating research components. The revised framework aligns transparency standards with these developments while maintaining flexibility across research domains.

More broadly, the revision reflects an evolution in how transparency standards are conceptualized and implemented. Rather than representing a departure from the original framework, TOP 2025 refines and extends it in response to experience gained through widespread adoption and application @grant.

```{r top-2025-table, echo=FALSE, results='asis'}
apa_table(top_2025, caption = "Research Practices")
```

### Problems and Barriers of the 2015 TOP Guidelines

Experience with TOP 2015 highlighted practical constraints affecting implementation at the journal level. Although many journals formally endorsed the guidelines, fewer adopted policies requiring enforcement or verification @toomey2025. In many cases, endorsement functioned as a public commitment to transparency without corresponding changes to editorial procedures @toomey2025.

These patterns reflected structural constraints within editorial workflows. Implementing transparency requirements involves not only policy revision but also procedures for evaluating compliance and managing exceptions. Survey research with editors indicated that limited staffing, competing responsibilities, and insufficient infrastructure often constrained implementation efforts @naaman2023. Without dedicated infrastructure or clearly defined operational standards, disclosure requirements were more readily implemented than policies requiring verification.

Ambiguities within the original framework further contributed to variability in implementation. Differences in how levels were defined across standards made consistent interpretation difficult, particularly in multidisciplinary journals @grant. Verification, in particular, lacked a uniform operational definition, limiting its feasibility as a broadly adopted editorial practice.

These experiences did not reflect opposition to transparency, but rather the complexity of translating general principles into operational editorial standards. The 2025 revision addresses these limitations by clarifying terminology, separating transparency from verification, and aligning the framework more closely with contemporary research and publication @centerforopenscience; @grant. In doing so, the updated guidelines provide a more precise foundation for journals seeking to implement and communicate transparency standards.

```{r top-comparison-table, echo=FALSE, results='asis'}
apa_table(top_comparison, caption = "Comparison of TOP Guidelines (2015 vs. 2025)")
```

## BRM’s Longstanding Commitment to Transparency

Since its founding, Behavior Research Methods has focused on publishing methodological contributions intended for implementation, evaluation, and reuse @brysbaert2021. The journal provides a venue for disseminating experimental paradigms, analytic procedures, software tools, and measurement approaches that support empirical research. These contributions function as shared methodological infrastructure, enabling researchers to apply, extend, and evaluate research procedures across contexts.

As methodological practices have evolved alongside advances in computing and data analysis, the journal has adapted its scope to support emerging forms of research infrastructure @brysbaert2021. Many articles provide not only conceptual descriptions but also executable implementations and detailed documentation, allowing methodological procedures to be examined and applied directly. This publication model supports the interpretability and evaluability of methodological claims by making the procedures underlying those claims accessible.

These practices developed prior to the formalization of transparency standards under initiatives such as the TOP Guidelines. The journal’s emphasis on methodological clarity, documentation, and accessibility reflects longstanding expectations that published methods should be sufficiently specified to permit independent implementation and evaluation @brysbaert2021. In this way, transparency has functioned as an implicit feature of the journal’s publication model.

The adoption of the TOP 2025 framework provides a standardized structure for articulating and formalizing these expectations @grant. By aligning editorial policies with the updated guidelines, BRM clarifies how transparency and verification practices apply within the journal’s methodological scope. This alignment supports consistent interpretation of transparency standards while maintaining continuity with the journal’s established publication practices.

The following section describes how BRM’s editorial policies correspond to the TOP 2025 framework and how these standards are applied in practice.

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
